# Robert Leigh
10 November 2017

This file is for data wrangling exercise 1, and involves modifying the learning2014 dataset
so that it is in a form which can be used for analysis. 


Read learning 2014 data using the read.table command and the web link. 
```{r}
lrn14 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS3-data.txt", sep="\t", 
                    header=TRUE)
```


Explore the structure of the data using str(). Most variables are integers, including those measured on a Likert (1:5) scale. Other integer variables not on the Likert scale include Age, Attitude, Points. gender is a factor with two levels (M/F).
```{r}
str(lrn14)
```

Explore the dimensions of the data using dim(). The data consists of 183 objects(rows) and 60 variables 
(columns).  
```{r}
dim(lrn14)
```


A new dataset with the variables gender, age, attitude, deep, stra, surf and points by combining questions in the learning2014 data. Combination variables were scaled by taking the mean, and observations where the exam points were zero were excluded. 

```{r}
library(dplyr)

lrn14$attitude <- lrn14$Attitude / 10

deep_questions <- c("D03", "D11", "D19", "D27", "D07", "D14", "D22", "D30","D06",  "D15", "D23", "D31")
surface_questions <- c("SU02","SU10","SU18","SU26", "SU05","SU13","SU21","SU29","SU08","SU16","SU24","SU32")
strategic_questions <- c("ST01","ST09","ST17","ST25","ST04","ST12","ST20","ST28")

deep_columns <- select(lrn14, one_of(deep_questions))
lrn14$deep <- rowMeans(deep_columns)
surface_columns <- select(lrn14, one_of(surface_questions))
lrn14$surf <- rowMeans(surface_columns)
strategic_columns <- select(lrn14, one_of(strategic_questions))
lrn14$stra <- rowMeans(strategic_columns)

keep_columns <- c("gender","Age","attitude", "deep", "stra", "surf", "Points")
learning2014 <- select(lrn14, one_of(keep_columns))


learning2014 <- filter(learning2014, Points > 0)

colnames(learning2014)[2] <- "age"
colnames(learning2014)[7] <- "points"
colnames(learning2014)

```


Check the structure of the modified dataset using str(). Now the dataset consists of only seven variables, gender, Age, attitude, deep, stra, surf and Points. Where means were taken, the variables are now numerical instead of integer values. Age is still an integer, and gender is still categorical. 

```{r}
str(learning2014)
```

Check the dimensions of the modified dataset using dim(). There are now 166 objects (rows), and 7 variables (columns) as expected. 
```{r}
dim(learning2014)
```


The new dataset was written to a .csv file for storage and further analysis using the write.csv() function. 

```{r}
write.csv(learning2014, file = "C:/HY-Data/LEIGH/course material/introduction_to_open_data_science/IODS_project_6_nov/data/learning2014.csv", row.names = FALSE)
```


Data was reimported from the .csv file to make sure everything worked properly. The structure and dimensions are the same as before writing to a .csv. 
```{r}
learning2014_import_check <- read.csv("C:/HY-Data/LEIGH/course material/introduction_to_open_data_science/IODS_project_6_nov/data/learning2014.csv")

str(learning2014_import_check)
dim(learning2014_import_check)
```






